Fail Faster

### 2025-03-21 paper/paper.tex
some words in intro from joe/bcp convo

Property-based testing (PBT) is a software-testing technique that uses
random test inputs to validate logical specifications. Recent studies on PBT usage~\cite{inpractice} in industry have shown that
practitioners often interleave PBT with their software development process, sometimes
testing their properties every time they save their code. This means that PBT is often given a very short time
window to find bugs---e.g., between 50 milliseconds and 30 seconds~\cite{inpractice}.
The faster it finds a failing test, the better.
% The goal of PBT is to fail tests as fast as possible. \jwc{First cut at ``fail fast''}
\iflater
\bcp{Feels like we're missing a trick in the transition here.  Can we
  say, ``Because performance is important, PBT libraries are already
  quite tuned.  [Can we substantiate that??]  But more is
  possible...'' }
\fi

@@

In summary, we show that monadic PBT generator DSLs incur significant performance costs across languages, and present a general technique for improving their efficiency while preserving their idiomatic abstractions. \jwc{needs a bit about RNG --- BCP thoughts?}
We make the following contributions:
\begin{enumerate}
    \item We identify two key sources of inefficiency in PBT generator libraries, which can significantly impact performance:
    abstraction overhead and \rand choice.
    \item We present \name{}, a technique based on staging that eliminates the abstraction overhead of generators. We implement \name{} in both OCaml and Scala 3, showcasing its cross-language applicability.
    \item We demonstrate that both \name{} and choosing a faster \rand yield substantial performance improvements to generator running time. Finally, we show
    that these performance improvements extend to significantly improved bug-finding speed. \jwc{This should be sell-ier: 2x speedups, etc. ``bug-finding speed'' isn't a race against time.}
\end{enumerate}

>>>>

Property-based testing (PBT) is a software-testing technique that uses
random test inputs to validate logical specifications. Recent studies on PBT usage~\cite{inpractice} in industry have shown that
practitioners often interleave PBT with their software development process, sometimes
testing their properties every time they save their code. This means that PBT is often given a very short time
window to find bugs---e.g., between 50 milliseconds and 30 seconds~\cite{inpractice}.
The faster it finds a failing test, the better.
% The goal of PBT is to fail tests as fast as possible. \jwc{First cut at ``fail fast''}
\iflater
\bcp{Feels like we're missing a trick in the transition here.  Can we
  say, ``Because performance is important, PBT libraries are already
  quite tuned.  [Can we substantiate that??]  But more is
  possible...'' }
\jwc{Agreed. Maybe one way to do this is to say someting in the next pagraph like ``much work has been poured building DSLs that let programers cleverly generate inputs to maximize the potential of finding bugs. Comparably less work has been poured into just making them fast.''}
\fi

@@

In summary, we show that monadic PBT generator DSLs incur significant performance costs across languages, and present a general technique for improving their efficiency while preserving their idiomatic abstractions. \jwc{needs a bit about RNG --- BCP thoughts?}
We make the following contributions:
\begin{enumerate}
    \item We identify two key sources of inefficiency in PBT generator libraries, which can significantly impact performance:
    abstraction overhead and \rand choice.
    \item We present \name{}, a technique based on staging that eliminates the abstraction overhead of generators. We implement \name{} in both OCaml and Scala 3, showcasing its cross-language applicability.
    \item We demonstrate that both \name{} and choosing a faster \rand yield substantial performance improvements to generator running time. Finally, we show
    that these performance improvements extend to significantly improved bug-finding speed. \jwc{This should be sell-ier: 2x speedups, etc. ``bug-finding speed'' isn't a race against time.}
    \jwc{We precisely quantify the performance effects of staging and faster rng across a range of testing scenarios --- i.e. a contrib is the fact that we can exactly pin down when one helps versus the other. The ``logically smarter'' your generator is, the more you get out of staging, the ``dumber'', the more you get out of rng.}
    \bcp{Agree with this. We need to beef up our claims here about what we learned about PBT-specifics.}
    \jwc{And this needs to be flowed through the rest of the paper.}
\end{enumerate}
